### 2020-04-29 steadystate.tex
Steady-state and transient chapters text without figures

This chapter discusses three main sets of simulation results. First, we
compare the key neutronics results between Serpent and Moltres for a static
model of the \gls{MSFR}, i.e. no salt flow, to assess the accuracy of the
six-group neutron diffusion model in Moltres on a fast-spectrum reactor. This
exercise also serves as a checking exercise for the group constant data
generated for Moltres from Serpent. Second, we present steady-state results
from Moltres with salt flow and verify it against published results, namely
from the Polimi/TUDelft and OpenFOAM models developed by Fiorina et al.
\cite{fiorina_modelling_2014} and Aufiero et al.
\cite{aufiero_development_2014}, respectively. Lastly, we present and compare
transient simulation results from Moltres with the aforementioned models.
Transient accident scenarios include an unprotected reactivity insertion, an
unprotected loss of flow, an unprotected loss of heat sink, a chilled inlet,
and a pump overspeed scenario.

\section{Static Model}

We calculated estimates of $k_{\text{eff}}$ values in Serpent and Moltres for
the \gls{MSFR} model with static salt (no flow), and uniform temperature
distributions. Moltres solves the six-group neutron diffusion equations as a
steady-state eigenvalue problem to find the $k_{\text{eff}}$ as it does not
currently have the capability to calculate the adjoint flux. Table
\ref{table:keff} shows the $k_{\text{eff}}$ values from Serpent and Moltres at
973 K and the corresponding salt density, while Table \ref{table:keff} shows
the $k_{\text{eff}}$ values for other temperatures at 100 K intervals. We
observe small discrepancies on the order of 100 pcm between the two codes
which we attribute to two main factors: the accuracy of the neutron diffusion
model, and the omission of the blanket tank structural material. The neutron
diffusion model is not as accurate as the other S$_{\text{N}}$ or
SP$_{\text{N}}$ deterministic methods nor the Monte Carlo approach in Serpent.
Regarding the omission of the blanket tank material, we replaced the
2 cm-thick structural material with blanket salt. This replacement may be
partly responsible for the higher $k_{\text{eff}}$ value calculated by Moltres
as the macroscopic fission cross sections of the blanket salt are non-zero for
the higher neutron energy groups. Nevertheless, the discrepancy is smaller
than the 228.5 pcm and 256.7 pcm discrepancies reported by Cervi et al.
\cite{cervi_development_2019} for their six-group $SP_3$ and neutron
diffusion methods, respectively, in OpenFOAM. The neutron diffusion model in
OpenFOAM is the same approach Aufiero et al. \cite{aufiero_development_2014}
used for their transient analysis of the \gls{MSFR}, albeit with one neutron
energy group.

\begin{table}[htb!]
    \small
	\centering
	\caption{$k_{\text{eff}}$ values from Serpent and Moltres at 973 K.}
	\begin{tabular}{l S c}
		\toprule
		{Code} & {$k_{\text{eff}}$} & {Difference wrt Serpent [pcm]}\\
		\midrule
		{Serpent} & 1.00662 \pm 0.00005 & {-}\\
		{Moltres with \glspl{DNP}} & 1.0079488 \pm 0.0000010 & 133\\
		{Moltres without \glspl{DNP}} & 1.0049369 \pm 0.0000010 & {-}\\
		\bottomrule
	\end{tabular}
	\label{table:keff}
\end{table}
%
\begin{table}[htb!]
    \small
	\centering
	\caption{$k_{\text{eff}}$ values from Serpent and Moltres at various
	temperatures from 800 K to 1400 K.}
	\begin{tabular}{S S S S}
		\toprule
		{Temperature [K]} & {$k_{\text{eff}}$ $\pm$ 1-$\sigma$ uncertainty
		(Serpent)} & {$k_{\text{eff}}$ (Moltres)} & {Difference wrt Serpent
		[pcm]}
		\\
		\midrule
		800  & 1.01996 \pm 0.00005 & 1.02117 & 121 \\
		900  & 1.01172 \pm 0.00005 & 1.01322 & 150 \\
		1000 & 1.00428 \pm 0.00005 & 1.00544 & 116 \\
		1100 & 0.99735 \pm 0.00005 & 0.99859 & 124 \\
		1200 & 0.99006 \pm 0.00005 & 0.99119 & 113 \\
		1300 & 0.98356 \pm 0.00005 & 0.98439 &  83 \\
		1400 & 0.97702 \pm 0.00005 & 0.97820 & 118 \\
		\bottomrule
	\end{tabular}
	\label{table:ktemp}
\end{table}

The absolute value of $k_{\text{eff}}$ largely impacts the steady-state
temperature of the reactor. We may adjust $k_{\text{eff}}$ by controlling
fissile inventory to raise or lower the operating temperatures and meet the
design specifications for inlet and outlet temperatures. For transient
analysis, the $\beta_{\text{eff}}$ and reactivity coefficients are arguably
more important as they dictate the duration and magnitude of the reactor
response to a transient initiator. The $\beta_{\text{eff}}$ values at 973 K,
shown in Table \ref{table:betaeff}, are in excellent agreement with a 2.89
pcm discrepancy.

\begin{table}[htb!]
	\centering
	\caption{$\beta_{\text{eff}}$ values from Serpent and Moltres at 973 K.}
	\begin{tabular}{l S S}
		\toprule
		{Code} & {$\beta_{\text{eff}}$ [pcm]} & {Difference wrt Serpent [pcm]}
		\\
		\midrule
		{Serpent} & 304.08 \pm .81 & {-}\\
		{Moltres} & 301.19 \pm .20 & 2.89\\
		\bottomrule
	\end{tabular}
	\label{table:betaeff}
\end{table}
%
\begin{figure}[htb!]
    \centering
    \includegraphics[width=.8\textwidth]{reactivity}
    \caption{Reactivity values from Serpent and Moltres. The Doppler
    reactivity values were calculated at a fixed density of 4.1249 g
    cm$^{-3}$. The density reactivity values were calculated at a fixed
    temperature of 973 K.}
    \label{fig:reactivity}
\end{figure}
%
\begin{table}[htb!]
	\centering
	\caption{Doppler, density, and total temperature coefficients
	for the temperature range of 800 K to 1400 K.}
	\begin{tabular}{l S S S S}
		\toprule
		{Code} & {$\alpha_D$ (log) [pcm]} & {$\alpha_D$ (linear) [pcm
		K$^{-1}$]} & {$\alpha_\rho$ [pcm K$^{-1}$]} & {$\alpha_T$ [pcm
		K$^{-1}$]} \\
		\midrule
		{Serpent} & -4034 \pm 14 & 3.737 \pm 0.013 & 3.424 \pm 0.013 & 7.165
		\pm 0.013 \\
		{Moltres} & {-} & {-} & {-} & 7.184\\
		\bottomrule
	\end{tabular}
	\label{table:alpha}
\end{table}

The temperature reactivity feedback arises mainly from Doppler broadening of
resonance absorption peaks and temperature-induced density changes. Although
Doppler coefficients typically show logarithmic dependence to temperature, we
report them, along with the other coefficients, as linear gradient
values of the reactivity given the relatively linear trend within the relevant
temperature range (Figure \ref{fig:reactivity}). Table \ref{table:alpha} shows
the temperature coefficients as described prior. The total temperature
coefficients from Serpent and Moltres show excellent agreement with a
discrepancy of 0.019 pcm K$^{-1}$.

Moltres also reproduced the six-group neutron spectrum very well from the
Serpent group constants. Figure \ref{fig:ntspec} compares the energy spectra
from Serpent and Moltres in the central fuel salt region. The six-group
neutron spectra overlap exactly over each other. More generally, the plot
shows the distinctive fast spectrum observed in the \gls{MSFR} with dips in
the spectrum corresponding to elastic scattering resonances from lithium and
fluorine. From this plot, we observe that the discrepancies in
$k_{\text{eff}}$ arise mainly from discretizing neutron energy into groups
rather than the neutron diffusion model itself. We could obtain a more
accurate representation of the neutronics in the \gls{MSFR} by using more
neutron energy groups but this would adversely impact simulation times in the
subsequent multiphysics finite element analyses.

\begin{figure}[htb!]
    \centering
    \includegraphics[width=.8\textwidth]{nt-spec}
    \caption{Reactivity values from Serpent and Moltres. The Doppler
    reactivity values were calculated at a fixed density of 4.1249 g
    cm$^{-3}$. The density reactivity values were calculated at a fixed
    temperature of 973 K.}
    \label{fig:ntspec}
\end{figure}

In summary, Moltres can replicate most of the relevant neutronics parameters
accurately with the group constant data from Serpent. While the
$k_{\text{eff}}$ values have discrepancies on the order of 100 pcm, they are
relatively small compared to results from other codes. Furthermore, the
$\beta_{\text{eff}}$ and temperature coefficients, which are important
parameters for modeling transient reactor behavior, are in excellent
agreement.

\section{Steady-State Results}

With the verification of Moltres' neutronics modeling capabilities in the
context of the \gls{MSFR}, we now move on to the multiphysics simulation
results. This section covers the steady-state multiphysics results from
Moltres.

@@

\subsection{Thermal-Hydraulics}

@@

\begin{figure}[htb!]
    \centering
    \begin{subfigure}[t]{.37\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.625\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-fiorina}
    \end{subfigure}
    \caption{Temperature and velocity fields in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models.}
    \label{fig:flow-temp}
\end{figure}

\begin{figure}[htb!]
    \centering
    \includegraphics[width=.6\textwidth]{flow}
    \caption{Fuel salt flow streamlines and velocity magnitude in the core.}
    \label{fig:flow}
\end{figure}

@@

\subsection{Neutronics}

\subsection{Decay Heat}

A decay heat model effectively redistributes a fraction of the heat source
from the center of the core to the entire loop. Thus, we expect to observe
a slight flattening of the temperature distribution across the entire primary
loop.

\section{Transient Results}

As noted by Fiorina et al. \cite{fiorina_modelling_2014}, explicit decay heat
modeling has a negligible effect on reactivity-, pump-, and enchanced
cooling-driven transients. The Polimi model presented results with the decay
heat model for the loss of heat sink transient scenario only. For a fair
comparison, we also

\subsection{Unprotected Reactivity Insertion}

\subsection{Unprotected Loss of Flow}

\subsection{Unprotected Pump Overspeed}

\subsection{Unprotected Loss of Heat Sink}

\subsection{Unprotected Chilled Inlet}

>>>>

With the verification of Moltres' neutronics modeling capabilities in the
context of the \gls{MSFR}, we now move on to the multiphysics simulation
results. This chapter covers the steady-state multiphysics simulation results
from Moltres.

@@

\section{Thermal-Hydraulics}

@@

\begin{figure}[htb!]
    \centering
    \begin{subfigure}[t]{.37\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp}
    \end{subfigure}
    \hfill
    \begin{subfigure}[t]{.625\textwidth}
        \centering
        \includegraphics[width=\textwidth]{flow-temp-fiorina}
    \end{subfigure}
    \caption{Temperature and velocity fields in the core from Moltres
    (left), Polimi (center), and TUDelft (right) models. The colors represent
    temperature according to the respective colorbars while the arrows
    represent velocity fields.}
    \label{fig:flow-temp}
\end{figure}

\begin{figure}[htb!]
    \centering
    \includegraphics[width=.6\textwidth]{flow}
    \caption{Fuel salt flow streamlines and velocity magnitude in the core.
    The colors represent velocity magnitude according to the colorbar on the
    right.}
    \label{fig:flow}
\end{figure}

@@

\section{Neutronics}

The neutron flux distribution represents the heat source distribution in a
nuclear reactor. Figure * shows the neutron flux distributions for all six
neutron energy groups. The distribution forms the symmetrical shape we expect
to see in a nuclear reactor with the peak at the center of the core. The
relatively lower temperatures near the center of the core enhances the neutron
flux peak but it is not a major concern as there are no structural parts
vulnerable to neutron damage in that region. The total flux at the center is
*, which is very close to values reported by Fiorina et al. 
\cite{fiorina_molten_2013} and Aufiero \cite{aufiero_development_2014}, albeit
their results are from their neutronics models at a fixed uniform temperature.

\begin{figure}[htb!]
    \centering
%    \includegraphics[width=.6\textwidth]{}
    \caption{Neutron flux distributions in the core for neutron energy groups
    1 to 6.}
    \label{fig:neutronflux}
\end{figure}

As mentioned before, the \glspl{DNP} are mobile in \glspl{MSR} and their
distributions do not directly correspond to the neutron flux distributions.
The location where delayed neutrons are emitted impacts their neutron
importance according to their proximity to fissile and parasitic isotopes.
Figure * shows the \gls{DNP} distributions for all eight \gls{DNP} groups. The
precursors from the shortest-lived group (Group 8) predominantly decay within
the core as their half-lives are shorter than the time it takes to reach the
outlet while the precursors from the longest-lived group (Group 1) are
effectively uniformly distributed due to their long half-lives relative to the
average circulation time. For the longer-lived groups, the mesh elements with
relatively high and low \gls{DNP} concentrations near the outlet and inlet,
respectively, suggest that the mesh is too coarse. Thus, we would recommend
further refinement for future work involving similar geometries.

\begin{figure}[htb!]
    \centering
%    \includegraphics[width=.6\textwidth]{}
    \caption{\gls{DNP} distributions in the core for \gls{DNP} groups
    1 to 8.}
    \label{fig:dnp}
\end{figure}

An important parameter for \gls{MSR} codes is the overall in-core
$\beta_{\text{eff}}$ value. This value represents the true effective delayed
neutron fraction in \glspl{MSR} that accounts for the loss of delayed neutrons
due to \gls{DNP} decay outside the core where fissions cannot be sustained. 
** DNP values and comparison **

\section{Decay Heat}

The inclusion of a decay heat model effectively redistributes a fraction of
the volumetric heat source from the center of the core to the entire loop.
Thus, we expect to observe a slight flattening of the temperature distribution
across the entire primary loop.
